WFH and broadband speed (title needs rework)

### 2020-12-22 paper/v1_sage/broadband.speed.covid.Rmd

<!-- Covid-19 and working from home, cities, urban structure -->
Furthermore, these multi-layered digital divides intersect with material divides
and the economic geography of the UK.
Following the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic],
different places have different industrial and occupational profiles,
and these affect the aggregated potential capacity of places for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south regions of England **replace 'UK'** regarding, for example, skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers some of the highest level of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
Not only are all three levels of digital divides associated, to a certain extent,
with or shaped by the geography of the UK, but the intersection of the digital and
material divides affects the capacity of places to overcome
some of the economic effects of the Covid-19 pandemic.
Importantly, this is the first time that digital technologies became an essential
tool for economic resilience for such a great part of the population.

@@

Meanwhile, those living in the largest cluster -- $6$, with $155$ LADs home to $29.5$
million people, experience aggregate mean upload speeds of about $1$Mb/s faster
than those in cluster $13$, but still lower than the other three large clusters
and most of the smaller clusters, suggesting a middling quality of service.
However, the time profile for cluster $6$ in Figure \ref{UpClusterL} shows
that upload speeds drop quickly from $6:00$ to $7:00$ on a Monday
and peak between $23:00$ and midnight on Wednesday and Thursday,
but tend to be lower during the working day.
Indeed, experienced mean upload speeds in the morning peak are $4.4$% lower
than in the evening peak -- a greater, more noticeable change than any of the other
large clusters experience, but smaller than any of the clusters with temporal
profiles shown in Figure \ref{UpClusterS}.
Clusters $8$ and $12$ also have mean upload speeds under $10$Mb/s, but higher than
clusters $1$ and $13$, and their standard deviation is not dissimilar.
However, this masks great variation in when lower speeds are experienced,
with the mean upload speeds much lower between $9:00-10:59$ than
between $19:00-20:59$ in cluster $8$, but slightly faster in the morning in
cluster $12$.

@@

Indeed cluster $12$ is the only cluster to experience higher speeds in the evening
peak, compared to the morning peak, suggesting that widespread telecommuting
has generally changed the temporal profile of internet activity throughout the UK.
Yet even if all but one cluster is showing slower speeds in the morning than
the evening, the reliability of internet services in different clusters
during the working day still varies considerably.
Interpreting this variation from the large spikes and dips shown on
Figures \ref{UpClusterL} and \ref{UpClusterS} is difficult,
but the statistics in Table \ref{up.cluster.descr} show that clusters $9$ and $11$
have the most reliable internet services.
The ratio of standard deviation to mean in both these clusters is below 55%,
and the ratio of upload speeds in the peak periods indicates that speeds are
only about 2% slower in the morning.
Mean speeds are also higher than in any other cluster, excluding cluster $2$,
where measures of reliability suggests poorer performance.

@@

In summary, the local authorities in clusters $9$ and $11$ experienced resilient
broadband internet that could support high levels of telecommuting.
Those in clusters $2$, $3$, $7$, and $10$ also experience higher than average
mean speeds and rank high to middle on measures of service reliability.
These LADs are *not* on the wrong side of the first level digital divide,
but how likely are they to be able to take advantage of their resilient ICT
infrastructure and services?
Meanwhile, cluster $6$ is not only the largest in terms of number of LADs
and population, it has the closest mean upload speed to the pre-clustered average
for the whole sample.
As well as average quality internet services, those in cluster $6$ also experience
average reliability for work purposes, ranking fifth behind the four other clusters
with populations over one million, but ahead of the smaller clusters.
Clusters $8$ and $12$ are also close to average mean upload speeds, but show very
different patterns in terms of reliability, whilst clusters $1$ and $13$ appear to
suffer most from a lack of quality internet services, with slow speeds and
high standard deviations.
With those in cluster $1$ in particular more likely to experience that poor
reliability during the morning peak, is this first level digital divide occurring
in areas where few are occupationally able to telecommute anyway, and what are the
implications for economic resilience?

@@

Using auxiliary regressions, we test whether the clusters that have higher mean
speeds and more reliable services consist of LADs that are more urban, affluent,
and / or more likely
to benefit from a choice of high quality internet services.
We also estimate which of our clusters are more likely to have a higher proportion of
occupations which enable telecommuting because of the nature of the work.
The results of these auxiliary regressions are presented in Table \ref{aux}.
The dependent variable is the LAD cluster membership as described in the
[methods and data](#sec:3) section
and equation \ref{eq1} and each column represent a different cluster.
The reference case is cluster $4$, which includes only the local authority of
Hambleton in North Yorkshire, a rural area of just over ninety thousand people.
Mean, experienced upload speeds in cluster $4$ are close to both the average speeds
for the $13$ clusters ($9.9$Mb/s) and the pre-clustered average for the whole sample
($9.3$Mb/s) as shown in \ref{up.cluster.descr}.
However, the standard deviation for cluster $4$ and the difference between average
speeds in the morning compared to the evening peak periods are indications of
worse reliability than many of the other clusters.
Hence, the results in Table \ref{aux} should be seen as relative
rather than absolute probabilities.

First, the number of speed tests run per cluster inhabitant between $9:00$-$10:59$
is an indication of satisfaction or at least a lack of concern over broadband speeds
and quality of service.
People in LADs assigned to clusters $6$ and $13$ were more likely to ran speed tests
(at a per capita basis)
at this time of day, whilst people in cluster $11$ were the least likely ones.
This high level of satisfaction matches the high quality and reliability of service
we would expect from the analysis in Section [4.1]](#sec:4.1).
Those in clusters $2$, $3$, $9$ and $10$ were also less likely to ran speed tests per
capita than those in the other six multi-authority clusters.
Cluster $7$ was the only cluster with higher mean speeds and middle to high measures
of reliability which was more likely to run more speed tests per capita.
This might be a reflection of the choice of ISP and connection options in those
LADs, as, along with cluster $12$, those in cluster $7$ were least likely to have
Virgin Media connections.
Fast Virgin Media cable connections have historically only been available
to $45$% of premises in the UK [@ofcom2016], where the more lucrative and
competitive market originally attracted the cable TV provider.
Those in clusters $2$, $3$, $9$ and $11$ benefit from a higher proportion of
Virgin connections, which is an indication that people in these clusters are
more likely to live in urban areas, with more choice of broadband services,
or in other words, are more likely to be on the right side of the first level,
infrastructure-based digital divide.

We employ distance from London and for the nearest metropolitan area (including
London) as two more variables depicting peripherality, urban structure and,
potentially, first level digital divides.
The broadband speed tests run in the authorities in cluster $3$ are more likely to
be taking place close to London than those run in any of the other clusters,
a result that makes sense considering two of the four authorities in cluster $3$
are the London commuter towns of Harlow and Luton.
However, even though London was also included in the variable calculating distance
from the centre of one of either the ten largest metropolitan areas in England,
Glasgow or Cardiff, tests run in cluster $3$ are likely to be furthest away.
Still, whilst significant, the effect is tiny and the other members of cluster $3$
are Corby and Eastbourne, two towns which, whilst not part of metropolitan areas,
are also not in remote parts of the country, and are home to an estimated population
of $175$ thousand people between them. **move maps to here?**
Meanwhile, LADs in cluster $7$ are most likely to be near the centre of a large
metropolitan area, even though the four local authorities of cluster $7$ include
no central urban boroughs, and only one LAD that is part of a metropolitan area of
governance - Tameside in Greater Manchester.
This may be why those in cluster $7$ are not served by Virgin Media.
It also is a demonstration of the complexity of experienced broadband upload speeds
as captured by time-series clustering, and their likely interaction with first level
digital divides.
Cluster $1$, which our analysis suggests lacks broadband resilience, contains five,
mainly rural authorities.
However, they are scattered around the country --
Ceredigion in West Wales, Darlington in the Northeast, Eden between the Lake District
and the North Penines, Rossendale in the Northwest, and Rutland in the East Midlands
-- and therefore, the results in Table \ref{aux} indicate that these authorities
are not as far from the centre of a metropolitan area as authorities in cluster $3$,
but furthest from London.

Internet resilience, as measured here, seems to be more nuanced than our --
arguably crude -- dummy variable depicting the North-South economic divide either,
which assigns $1$ to LADs locates in Greater London, Southeast, Southwest and East
of England regions.
The authorities most likely to be in the South are those in cluster $13$,
which was identified as having the slowest mean upload speeds of any of the clusters,
and thus a low level of service.
However, cluster $13$ does include some rural, remote areas of the country,
such as Northwest Scotland, Cornwall and Powys in Wales as shown on Figure
\ref{map.up.clusters}.
It also includes the major metropolitan centres of Bristol, Liverpool,
and Newcastle, as well as nine (of $32$) London Boroughs and plenty of home county
and suburban areas.
Although by the standard deviation measure cluster $13$ has unreliable internet
services, speed variation was less during the morning peak.
More broadly, the estimates for this variable are not consistent between the
two clubs of LAD clusters: those who are in the wrong and the right side of
digital divides.

Regarding the rest of the socio-economic variables we see a varied
picture. Clusters $9$ and $11$, which are the ones benefiting by the most resilient
internet connections are likely to consist of LADs with higher share of managerial
and highly skilled occupations as well as higher average earnings. Interestingly,
both clusters are characterised by lower share of occupations that are not usually
associated with teleworking (professional, associated professional and technical,
administrative, leisure and machine operation jobs), but also by lower number of
business per habitant and. However, the likelihood of hosting highly skilled
residents is not consistent with cluster $11$ having a negative coefficient for the
NVQ4 variable. This might be an indication of the higher returns that resilient
broadband infrastructure generates for cluster $9$ versus cluster $11$. While both
of these cluster consist more likely by large LADs in terms of population, job
density if higher for LADs in cluster $11$ indicating more labour intensive
economic structures contrary to LADs in cluster $9$.

A similar divided picture can be built for the next tier of clusters: $2$, $3$,
$7$, $10$. LADs in these clusters experience higher than average mean speeds and
rank high to middle on measures of service reliability. As mentioned before, LADs
in these clusters are on the right side of the first level digital divide. However,
while all of them are more likely to be populated by people employed in managerial
jobs, only LADs in cluster $2$ and $10$ are more likely to have highly skilled
work force, which can more easily take advantage of working from home opportunities.
A more mixed picture can be built for LADs in clusters $3$ and $7$, which are
associated, on the one hand, with lower skills and, on the other hand, with higher
likelihood of individuals in managerial occupations. A similar contradicting
pattern can be illustrated for cluster $6$, which is the largest one with average 
quality of internet service. Arguably, LADs in clusters $2$ and $10$ are the ones
benefiting more by resilient internet infrastructure in an era that working
from home became a vehicle for economic resilience.

Moving on to the next tier of clusters $8$ and $12$, which are characterised by
LADs enjoying close to average mean upload speeds, but different patterns of internet
reliability, we see a differentiated picture. On the one hand, both of them are
likely to host individuals employed in managerial occupations while the opposite
applies on skilled, professional and leisure occupations. On the other hand,
LADs in cluster $8$ are characterised by a very small positive and LADs in cluster
$12$ by a negative likelihood to host highly skilled individuals. Moreover, LADs
in these two clusters have a rather
similar profile regarding the occupational variables, but different tendencies
regarding the stock of businesses. In total, LADs in these clusters do not
benefit by great internet services and could only marginally benefit by working
from home opportunities.

Lastly, clusters $1$ and $13$, which appear to suffer most from a lack of quality
internet services are more likely to also suffer from low skilled workforce. LADs
in cluster $13$. Contradictory, the likelihood of workforce employed in managerial
occupations is positive for LADs in both clusters, but only individuals in LADs
in cluster $13$ are more likely to enjoy high earnings.

@@

those in cluster $6$, including people living in Birmingham, Leeds and
Sheffield, in twelve London Boroughs, and many suburban areas and smaller cities like
Cardiff, Oxford and Cambridge
This lower level of reliability may be as a result of increased demand,
e.g. for telecommuting, especially as there are few truly remote areas
in this cluster.

The eight LADs in clusters $3$ and $10$ are home to over one million people,
and are all more urban and suburban than cluster $1$, including the London Borough
of Newham, the Southeast commuter towns of Harlow and Luton, the Birmingham suburbs
of Bromsgrove and Cannock Chase, Dundee City, and the towns of Eastbourne and Corby.

Although not all of the local authorities in these clusters $9$ and $11$ are urban,
among the $65$ LADs are Manchester and three of the nine other boroughs of
Greater Manchester, Glasgow, Nottingham, both Portsmouth and Southampton,
eight London Boroughs, four of the seven constituent authorities of the
West Midlands Combined Authority.
There are also a number of other tightly bounded urban areas, such as Aberdeen,
Blackpool, Ipswich, Norwich, Slough, and Stevenage, and urban areas at the centre
of less confined districts like Burnley, Milton Keynes, and Northampton.

>>>>

<!-- Covid-19 and working from home, cities, urban structure -->
Furthermore, these multi-layered digital divides intersect with material divides
and the economic geography of the UK.
Following the regional economic resilience literature, which
underlines the differentiated capacity of cities and regions to escape or
recover from economic crises [@martin2012regional; @kitsos2018economic],
different places have different industrial and occupational profiles,
and these affect the aggregated potential capacity of places for telecommuting.
Such profiles are associated with longstanding inequalities in the UK and their
spatial representation as a North-South divide [@martin_north_south].
Various studies have illustrated severe inequalities between the north and the
south regions of England regarding, for example, skills and human capital,
unemployment, productivity and prosperity [@lee2014grim; @mccann2020perceptions;
@dorling2018peak].
Some scholars have even argued that the UK suffers some of the highest level of
interregional inequalities in the global north [@gal2018reducing; @mccann2016uk].
Not only are all three levels of digital divides associated, to a certain extent,
with or shaped by the geography of the UK, but the intersection of the digital and
material divides affects the capacity of places to overcome
some of the economic effects of the Covid-19 pandemic.
Importantly, this is the first time that digital technologies became an essential
tool for economic resilience for such a great part of the population.

@@

Meanwhile, those living in the largest cluster -- $6$, with $155$ LADs home to $29.5$
million people, experience aggregate mean upload speeds of about $1$Mb/s faster
than those in cluster $13$, but still lower than the other three large clusters
and most of the smaller clusters, suggesting a middling quality of service.
However, the time profile for cluster $6$ in Figure \ref{UpClusterL} shows
that upload speeds are highest at $6:00$ on a Monday morning
and peak between $23:00$ and midnight on Wednesday and Thursday,
but tend to be lower during the working day.
Indeed, experienced mean upload speeds in the morning peak are $4.4$% lower
than in the evening peak -- a greater, more noticeable change than any of the other
large clusters experience, but smaller than any of the clusters with temporal
profiles shown in Figure \ref{UpClusterS}.
Clusters $8$ and $12$ also have mean upload speeds under $10$Mb/s, but higher than
clusters $1$ and $13$, and their standard deviation is not dissimilar.
However, this masks great variation in when lower speeds are experienced,
with the mean upload speeds much lower between $9:00-10:59$ than
between $19:00-20:59$ in cluster $8$, but slightly faster in the morning in
cluster $12$.

@@

Indeed cluster $12$ is the only cluster to experience higher speeds in the evening
peak, compared to the morning peak, suggesting that widespread telecommuting
has generally changed the temporal profile of internet activity throughout the UK.
Yet even if all but one cluster is showing slower speeds in the morning than
the evening, the reliability of internet services in different clusters
during the working day still varies considerably.
Interpreting this variation from the large spikes and dips shown on
Figures \ref{UpClusterL} and \ref{UpClusterS} is difficult,
but the statistics in Table \ref{up.cluster.descr} show that clusters $9$ and $11$
have the most reliable internet services.
The ratio of standard deviation to mean in both these clusters is below 55%,
and the ratio of upload speeds in the peak periods indicates that speeds are
only about 2% slower in the morning than the evening.
Mean speeds are also higher than in any other cluster, excluding cluster $2$,
where measures of reliability suggests poorer performance.

@@

In summary, the local authorities in clusters $9$ and $11$ experienced resilient
broadband internet that could support high levels of telecommuting.
Those in clusters $2$, $3$, $7$, and $10$ also experience higher than average
mean speeds and rank middle to high on measures of service reliability.
These LADs are *not* on the wrong side of the first level digital divide,
but how likely are they to be able to take advantage of their resilient ICT
infrastructure and services?
Meanwhile, cluster $6$ is not only the largest in terms of number of LADs
and population, it has the closest mean upload speed to the pre-clustered average
for the whole sample.
As well as average quality internet services, those in cluster $6$ also experience
average reliability for work purposes, ranking fifth behind the four other clusters
with populations over one million, but ahead of the smaller clusters.
Clusters $8$ and $12$ are also close to average mean upload speeds, but show very
different patterns in terms of reliability, whilst clusters $1$ and $13$ appear to
suffer most from a lack of quality internet services, with slow speeds and
high standard deviations.
With those in cluster $1$ in particular more likely to experience that poor
reliability during the morning peak, is this first level digital divide occurring
in areas where few are occupationally able to telecommute anyway, and what are the
implications for economic resilience?

@@

Using auxiliary regressions, we test whether the clusters that have higher mean
speeds and more reliable services consist of LADs that are more urban, affluent,
and / or more likely
to benefit from a choice of high quality internet services.
We also estimate which of our clusters are more likely to have a higher proportion of
occupations which enable telecommuting because of the nature of the work.
The results of these auxiliary regressions are presented in Table \ref{aux}.
The dependent variable is the LAD cluster membership as described in the
[methods and data](#sec:3) section
and equation \ref{eq1} and each column represent a different cluster.
The reference case is cluster $4$, which includes only the local authority of
Hambleton in North Yorkshire, a rural area of just over ninety thousand people.
Mean, experienced upload speeds in cluster $4$ are close to both the average speeds
for the $13$ clusters ($9.9$Mb/s) and the pre-clustered average for the whole sample
($9.3$Mb/s) as shown in Table \ref{up.cluster.descr}.
However, the standard deviation for cluster $4$ and the difference between average
speeds in the morning compared to the evening peak periods are indications of
worse reliability than many of the other clusters.
Hence, the results in Table \ref{aux} should be seen as relative
rather than absolute probabilities.

First, the number of speed tests run per cluster inhabitant between $9:00$-$10:59$
is an indication of satisfaction or at least a lack of concern over broadband speeds
and quality of service.
People in LADs assigned to clusters $6$ and $13$ were more likely to run speed tests
(at a per capita basis)
at this time of day, whilst people in cluster $11$ were the least likely to do so.
This high level of satisfaction matches the high quality and reliability of service
we would expect from the analysis in Section [4.1]](#sec:4.1).
Those in clusters $2$, $3$, $9$ and $10$ were also less likely to ran speed tests per
capita than those in the other six multi-authority clusters.
Cluster $7$ was the only cluster with higher mean speeds and middle to high measures
of reliability which was more likely to run more speed tests per capita.
This might be a reflection of the choice of ISP and connection options in those
LADs, as, along with cluster $12$, those in cluster $7$ were least likely to have
Virgin Media connections.
Fast Virgin Media cable connections have historically only been available
to $45$% of premises in the UK [@ofcom2016], where the more lucrative and
competitive market originally attracted the cable TV provider.
Those in clusters $2$, $3$, $9$ and $11$ benefit from a higher proportion of
Virgin connections, which is an indication that people in these clusters are
more likely to live in urban areas, with more choice of broadband services,
or in other words, are more likely to be on the right side of the first level,
infrastructure-based digital divide.

We employ distance from London and for the nearest metropolitan area (including
London) as two more variables depicting peripherality, urban structure and,
potentially, first level digital divides.
The broadband speed tests run in the authorities in cluster $3$ are more likely to
be taking place close to London than those run in any of the other clusters,
a result that makes sense considering two of the four authorities in cluster $3$
are the London commuter towns of Harlow and Luton.
However, even though London was also included in the variable calculating distance
from the centre of one of either the ten largest metropolitan areas in England,
or Glasgow or Cardiff, tests run in cluster $3$ are likely to be furthest away.
Still, whilst significant, the effect is tiny and the other members of cluster $3$
are Corby and Eastbourne, two towns which, whilst not part of metropolitan areas,
are also not in remote parts of the country, and are home to an estimated population
of $175$ thousand people between them. **move maps to here?**

Meanwhile, LADs in cluster $7$ are most likely to be near the centre of a large
metropolitan area, even though the four local authorities of cluster $7$ include
no central urban boroughs and only one LAD that is part of a metropolitan area of
governance - Tameside in Greater Manchester.
This may be why those in cluster $7$ are not served by Virgin Media.
It also is a demonstration of the complexity of experienced broadband upload speeds
as captured by time-series clustering, and their likely interaction with first level
digital divides.
Cluster $1$, which our analysis suggests lacks broadband resilience, contains five,
mainly rural authorities.
However, they are scattered around the country --
Ceredigion in West Wales, Darlington in the Northeast, Eden between the Lake District
and the North Penines, Rossendale in the Northwest, and Rutland in the East Midlands
-- and therefore, the results in Table \ref{aux} indicate that these authorities
are not as far from the centre of a metropolitan area as authorities in cluster $3$,
but furthest from London.

Internet resilience, as measured here, seems to be more nuanced than our --
arguably crude -- dummy variable depicting the North-South economic divide either,
which assigns $1$ to LADs locates in Greater London, Southeast, Southwest and East
of England regions, and $0$ to the rest.
The authorities most likely to be in the South are those in cluster $13$,
which was identified as having the slowest mean upload speeds of any of the clusters,
and thus a low level of service.
However, cluster $13$ does include some rural, remote areas of the country,
such as Northwest Scotland, Cornwall and Powys in Wales as shown on Figure
\ref{map.up.clusters}.
It also includes the major metropolitan centres of Bristol, Liverpool,
and Newcastle, as well as nine (of $32$) London Boroughs, 
including the City of London and plenty of home county and suburban areas.
Although by the standard deviation measure cluster $13$ has unreliable internet
services, speed variation was less during the morning peak, suggesting that 
the estimates for reliability are inconsistent. 
Considering that this is one of the largest clusters, and thus the averages 
incorporate more noise than some of the smaller clusters, 
it may be that the LADs in this cluster do not all suffer equally from 
a first level digital divide.

Yet we need to consider the results for other variables in Table \ref{aux} to better
determine whether clusters $1$ and $13$, which appear to suffer most from a lack 
of quality internet services, are also more likely to have a low skilled workforce, 
less able to benefit from telecommuting. 
Cluster $1$ has the lowest propotion of educated people of any of the clusters,
and the lowest earnings, despite recording the highest proportion of managerial, 
professional jobs, and the second highest proportion of tech jobs.
Rural areas such as those in cluster $1$ are often home to many retired people,
which might explain these results or perhaps, as these figures are relative,
we could note that cluster $1$ also has a greater proportion of skilled trades
than other clusters.
Either way, it seems likely that the first level digital divide reinforces 
other inequities in cluster $1$.
In comparison, those cluster $13$ are more likely to earn more and have a better
education.
Cluster $13$ also has the most businesses per inhabitant, but is somewhere in the 
middle in terms of job density.
If this is an indication of a high number of SMEs, it might give some explanation
for the variable internet quality, considering that small businesses have not been
seen as the most valuable customers for higher speed broadband services **insert ref**.

Cluster $6$ is the largest cluster, and thus, like cluster $13$, there is more 
noise within the averages we use to measure internet quality and reliability,
and the result indicates an average level of service, even if better than LADs
in cluster $13$.
Cluster $6$ also falls towards the middle of the clusters on many of the socio-
economic variables in Table \ref{aux}. 
It ranks third or fourth out of the eleven clusters in the table in terms of
having a higher proportion of managerial, professional, and tech jobs, as well as 
higher earnings, but is fourth from bottom in terms of education.
The LADs in cluster $6$ are also diverse, with few truly remote areas, 
but many urban areas in both the North, including Birmingham, Leeds and Sheffield, 
and South, including twelve London Boroughs, and many suburban areas and 
smaller cities like Oxford and Cambridge.
The capital cities of the other UK nations, Belfast, Cardiff and Edinburgh, are
also in this cluster, suggesting perhaps that the lower level of reliability 
discussed in Section [4.1](#sec:4.1) may be as a result of increased demand,
e.g. for telecommuting.

In comparison, clusters $9$ and $11$, which benefit from the most resilient
internet connections, are characterised by lower share of occupations that are  
may benefit from the ability to telecommute (e.g. managerial and technical). 
Cluster $9$ is more highly educated and higher earning than cluster $11$, 
has fewer businesses per inhabitant and less job density.
Indeed, cluster $11$ has a negative coefficient for the NVQ4 variable, whilst
for cluster $9$ the coefficient is positive, and cluster $11$ has the second
highest job density and cluster $9$ the second lowest. 
The former might be an indication of the higher returns that resilient
broadband infrastructure generates for cluster $9$ versus cluster $11$. 
Indeed, we did detect slightly more slowdown in the AM peak in cluster $9$ than in
cluster $11$ (see Table \ref{up.cluster.descr}).
Meanwhile, a propensity for a high job density might indicate more labour intensive 
economic structures in cluster $11$ than in the LADs in cluster $9$.

The LADs in clusters $9$ and $11$ also have opposite signs for their likelihood 
of being in the South of England.
Those in cluster $9$ less likely to be in the South and those in cluster $11$ 
more likely, although the coefficient for cluster $9$ might be skewed by 
the presence of the Scottish cities of Glasgow and Aberdeen, 
as Scotland has a different economic profile than England.
Still, both clusters consist of larger LADs in terms of population, 
including the urban centre of Manchester and three of the nine other districts of
Greater Manchester, Nottingham, both Portsmouth and Southampton,
eight London Boroughs, and four of the seven constituent authorities of the
West Midlands Combined Authority.
Among the $65$ LADs are also a number of other tightly bounded urban areas, 
such as Blackpool, Ipswich, Norwich, Slough, and Stevenage, and urban areas 
at the centre of less confined districts like Burnley, 
the digital hub of Milton Keynes, and Northampton.
Thus, our findings confirm the importance of geography and urban locations 
for being on the right side of the first level digital divide, and suggest that 
internet resilience cuts across other digital and economic divides.

A similarly divided picture can be built for the next tier of clusters: $2$, $3$,
$7$, $10$. LADs in these clusters experience higher than average mean speeds and
rank high to middle on measures of service reliability. 
LADs in these clusters are on the right side of the first level digital divide. 
However, only LADs in cluster $2$ and $10$ are more likely to have highly skilled
work force, which might more easily take advantage of working from home 
opportunities.
Cluster $2$ is comprised of just two LADs, with the lowest job density of any
cluster -- these are East Lothian near Edinburgh, and North East Lincolnshire, 
home to Grimsby.
Cluster $10$ is comprised of four LADs, including two peripheral suburban areas 
of Birmingham, the East London Borough of Newham, and Dundee.
Suburbs are considered the most likely urban form in which telecommuters live, 
and Dundee has a reputation for tech startups **insert ref**.
Yet although those in clusters $2$ and $10$ may be making the most of their 
resilient internet, those in cluster $7$ have the highest earnings.
The four LADs in cluster $7$ include the suburbs South of Belfast, a suburban 
district of Greater Manchester, the villages between York and Leeds, and a
couple of small West Midlands towns.
LADs in cluster $3$ are associated more with lower skills and have the highest
proportion of individuals in machine operation jobs. 
They include the Southeast commuter towns of Harlow and Luton and the slightly more
distant towns of Eastbourne and Corby.
Arguably, LADs in clusters $3$ benefit the least from their resilient internet 
infrastructure in an era when working from home became a vehicle for economic 
resilience.

Finally, clusters $8$ and $12$, are characterised by LADs enjoying close to average 
mean upload speeds, but different patterns of internet reliability. 
LADs in cluster $8$ are likely to host more individuals employed in tech and
administrative occupations than any other cluster, as well as many in managerial 
occupations, whilst the opposite applies to skilled trade and leisure occupations. 
These LADs are characterised by a very small positive likelihood of more educated 
residents, but are not significantly likely to earn more than other clusters.
Including suburbs of Leicester, Cardiff, Birmingham, the green belt area south of 
London straddling the M25 motorway, and rural West Devon, the LADs in cluster $8$
seem to likely to have the skills and occupations that would benefit from 
quality internet services, but suffer from poor internet resilience and reliability.

In comparison, LADs in cluster $12$ are less likely to be able to benefit from
quality internet services if they had them, with fewer individuals achieving 
NVQ4 or better and lower levels of occupations that would benefit from homeworking. 
Yet LADs in this cluster benefit from the second highest level of earnings, and
have the largest stock of businesses per inhabitant and the highest job density. 
This density of businesses could be why cluster $12$ is the only cluster with higher
speeds in the morning peak during the study period.
If people are at home, business premises might well be abandoned.
Cluster $12$ is home to $1.5$million people spread across twelve LADs, including
the City of Westminster and the London Borough of Hammersmith.
However, it also includes a LAD in Northern Ireland, two in Wales, two suburban areas
North of London, Preston in the Northwest, East Staffordshire in the Midlands, and
Fenland in the East of England.
This spatial diversity demonstrates that the temporal clustering of internet 
resilience does not create spatial clusters, and digital divides do not necessarily
overlap with economic ones.

@@
